Figures to insert in text

### 2018-04-03 Chapters/Chapter02.tex
theorem form

Each parameter set choice like this will lead to particular numbers --- many of them emergent---, such as the access radius size, critical distance, memory capacity, and so forth.

One intriguing question here is:  is there a `better' number of dimensions and hard locations?  If so, can such numbers be better studied algebraically or numerically?

How should these parameters be compared?  What are the tradeoffs that should be considered?  What are the `best' benchmarks possible?

@@

A hypercube with n dimensions can be divided by two hypercubes with $n-1$ dimensions. Is there an algorithm that separates the area of each hard location in such a form that there exists a function mapping each bitstring in $\{0,1\}^n$ to the set of hard locations it `belongs to'?  Though this would break Kanerva's assumption of a randomly yet uniformly distributed set of hard locations --- for a perfectly symmetrical set of hard locations ---, there could be large performance gains if such a mapping function from a bitstring to its corresponding set of nearest hard locations exists. Note that others have attempted heuristic approaches to this\footnote{\citet{Kanerva2009} has recently mentioned that `The computationally most efficient implementation of it, by \citet{Karlsson95afast}, is equivalent to the RAM-based WISARD of  \citet{aleksander1982computer}.' Note, however, that Karlsson's model seems to require $O(2^n)$ hard-locations, and may introduce non-uniform asymmetries in the space.}.

@@

\section{Illuminating and paving a pathway}

@@

Of course, for the skeptical reader who may consider software a less worthy pursuit, there is also new work here.  The mathematics of the model has been shown to be correct numerically (with a single, small, anomaly); we have shown how to execute unsupervised learning with nothing besides operations original to the SDM; we have studied the generalized Murilo read; we have seen noise filtering; the death of neurons; how information-theory may be of use; and finally, we have reproduced numerous of the original propositions put forth by Kanerva.  The emphasis might have been on the \emph{breadth of topics}, in detriment of depth here or there.  But this is due to our research group's enthusiasm for the topic; we do indeed believe that SDM is --- if not correct --- extremely close to a full scientific understanding of human long-term memory.  If so, it is such a monumental achievement that we want readers to be able to see all of what we see and imagine the vastness of possibilities.  The work on, say, reinforcement learning, is most definitely not the definitive work we will see on the subject, but a challenge left for readers to contemplate. Ralph Waldo Emerson once said \emph{do not go where the path may lead. Go, instead, where there is no path, and leave a trail.}  Professor Pentti Kanerva has left the trail.  It is my job to illuminate it and to pave it and to clear it; to try to deliver an easier pathway for the next generation.  Some essays completely shut the door close at the end; this one intends to leave it wide open. As the reader might have noticed, this final section does not read as an analysis of the work done; it reads, instead, as a \emph{desideratum}, a prologue, a yearning for others to join me in imagining the shape of things to come.

% TODO Include in the to do list simulations based on the suggestion of Murilinho.

>>>>

Each parameter set choice like this will lead to particular numbers --- many of them emergent ---, such as the access radius size, critical distance, memory capacity, and so forth.  For example, the number of dimensions $n$ leads to the number of standard deviations available on the space: $2 \sqrt{n}$.  One then must choose a proper multiple $m\times\sigma$: Kanerva (and most in the literature) have been using $m=3$, as $\mu-3\sigma$ selects approximately $1/1000$ of the space. These constants will determine the size of the access radius.  Then one must choose the number of hard locations, which will determine how many hard-locations are activated on average; and so on and so forth...

One intriguing question here is:  is there a `better' number of dimensions and hard locations?  If so, can such numbers be better studied algebraically or numerically? How should these parameters be compared?  What are the tradeoffs that should be considered?  What are the `best' benchmarks possible?

@@

A hypercube with n dimensions can be divided by two hypercubes with $n-1$ dimensions. Is there an algorithm that separates the area of each hard location in such a form that there exists a function mapping each bitstring in $\{0,1\}^n$ to the set of hard locations it `belongs to'?  Though this would break Kanerva's assumption of a randomly yet uniformly distributed set of hard locations --- for a perfectly symmetrical set of hard locations ---, there could be large performance gains if such a mapping function from a bitstring to its corresponding set of nearest hard locations exists. Note that others have attempted heuristic approaches to this\footnote{\citet{Kanerva2009} has recently mentioned that `The computationally most efficient implementation of it, by \citet{Karlsson95afast}, is equivalent to the RAM-based WISARD of  \citet{aleksander1982computer}.' Note, however, that Karlsson's model may introduce non-uniform asymmetries in the space, and seems to require a staggering $O(2^n)$ hard-locations.}.

@@

\section{Illuminating, paving, clearing}

@@

Of course, for the skeptical reader who may consider software a less worthy pursuit, there is also new work here.  The mathematics of the model has been shown to be correct numerically (with a single, small, anomaly); we have shown how to execute unsupervised learning with nothing besides operations original to the SDM; we have studied the generalized Murilo read; we have seen noise filtering; the death of neurons; how information-theory may be of use; and finally, we have reproduced numerous of the original propositions put forth by Kanerva.  The emphasis might have been on the \emph{breadth of topics}, in detriment of depth here or there.  The work on, say, reinforcement learning, is most definitely not the last word we will see on the subject, but a challenge left for readers to contemplate. But this is due to our research group's enthusiasm for the topic; we do indeed believe that SDM is --- if not correct --- extremely close to a full scientific understanding of human long-term memory.  If so, it is such a monumental achievement that we want readers to be able to see all of what we see and imagine the vastness of possibilities.

Ralph Waldo Emerson once said: \emph{do not go where the path may lead. Go, instead, where there is no path, and leave a trail.}  Professor Pentti Kanerva has left the trail.  It is my job to illuminate it and to pave it and to clear it; to try to deliver an easier pathway for the next generation.  Some essays completely shut the door close at the end; this one intends to leave it wide open. As the reader might have noticed, this final section does not read as an analysis of the work done; it reads, instead, as a \emph{desideratum}, a prologue, a yearning for others to join me in imagining the shape of things to come.

% TODO

### 2018-04-03 Chapters/sdm-critical-distance.tex
theorem form

This is an important insight: a hard location is activated because it shares many dimensions with the items read from or written onto it. Imagine the `counter's eye view':  each individual counter `likes' to write its own corresponding bit-address more than it likes the opposite; as each hard-location has a say in its own area (and nowhere else).

>>>>

This is an important insight: a hard location is activated because it shares many dimensions with the items read from or written onto it. Imagine the `counter's eye view':  each individual counter `likes' to write on its own corresponding bit-address more than it likes the opposite; as each hard-location has a say in its own area (and nowhere else).

### 2018-04-03 kanerva-answer.tex
theorem form

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{subfig}

@@

[Thesis text....] ...As one can easily notice, we have got exactly the same curve as Kanerva. Both his and our model expect that, after reading, say, from 550 bits of distance from a written bitstring, we should obtain the expected $n/2$ equator distance. This question has intrigued us, and here we look for a more analytic explanation than merely interference from the other written attractors. Let us turn back to mathematics to study this anomaly.

@@

   "Our results show that the theoretical prediction is not accurate. There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond ~500 bits
   (Figure \ref{fig:sdm-10000w-table-7-2})." \\

@@

So what is the probability of shared bit-values given that we know the access radius $r$ between the address and a hard-location?  Applying the law of total probability:

\begin{align}
P(x_i = y_i | d(x, y) \le r) = \sum_{k=0}^{r} P(x_i = y_i | d(x, y) = k \le r) \cdot P(d(x, y) = k | d(x, y) \le r)
\end{align}

@@

Finally, as $\frac{n-k}{n} \binom{n}{k} = \binom{n-1}{k}$,

\begin{align}
P(x_i = y_i | d(x, y) \le r) = \frac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}}
\end{align}

>>>>

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{amsthm}

\newtheorem{theorem}{Theorem}[section]
\usepackage{etoolbox}
\AtBeginEnvironment{theorem}{\large}

@@

[Thesis text....] ...As one can easily notice, we have got exactly the same curve as Kanerva. Both his and our model expect that, after reading, say, from 550 bits of distance from a written bitstring, we should obtain the $n/2$ equator distance. We have not, however.  This question has intrigued us, and here we look for a more analytic explanation than merely interference from the other written attractors. Let us turn back to mathematics to study this anomaly.

@@

   ``Our results show that the theoretical prediction is not accurate. There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond ~500 bits
   (Figure \ref{fig:sdm-10000w-table-7-2}).'' \\

@@

So what is the probability of shared bit-values given that we know the access radius $r$ between the address and a hard-location?
\bigskip

\begin{theorem}
%\emph{(Lagrange's Theorem)}
\label{T1}
$P(x_i = y_i | d(x, y) \le r) = \frac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}}$
\end{theorem}

\begin{proof}
    Applying the law of total probability to the left-hand expression we obtain

    \begin{align}
    \sum_{k=0}^{r} P(x_i = y_i | d(x, y) = k \le r) P(d(x, y) = k | d(x, y) \le r)
    \end{align}

@@

    Finally, the combinatorial identity

    \begin{align}
    \frac{n-k}{n} \binom{n}{k} = (n-k)\binom{n-1}{k} =(n-k)\binom{n-1}{n-k-1} = \binom{n-1}{k}
    \end{align}

    closes the theorem.

\qed
\end{proof}
